---
title: "Project 1"
author: "Grace Fischer"
date: "09/15/26"
format: html
---
```{r}
#| label: access-data
#| warning: false

## Install the tidyverse package
if (!require("tidyverse", quietly = TRUE)) {
    install.packages("tidyverse")
}

library("tidyverse")

## Install the tidytuesdayR package
if (!require("tidytuesdayR", quietly = TRUE)) {
    install.packages("tidytuesdayR")
}

library("here")
library("tidyverse")

# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here("data", "chocolate.RDS"))) {
    url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-01-18/chocolate.csv"
    chocolate <- readr::read_csv(url_csv)

    # save the file to RDS objects
    saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}

# Read dataset locally from computing enviorment
chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)

# Glimpse at the data 
glimpse(chocolate)
```


Part 1
```{r}
#| label: Part 1: 1. Explore Data - Histogram
#| warning: false

## Histogram where bin num = 10 
hist(chocolate$rating, main = "Overall distribution of rating scores (bin = 10)", xlab =  "Rating", ylab = "Frequency", breaks = 10)

## Histogram where bin num = 15 
hist(chocolate$rating, main = "Overall distribution of rating scores (bin = 15)", xlab =  "Rating", ylab = "Frequency", breaks = 15)

## Histogram where bin num = 20 (breaks this way because r typically favors pretty numbers over round, length out is because it produces n-1 bins)
hist(chocolate$rating, main = "Overall distribution of rating scores (bin = 20)", xlab =  "Rating", ylab = "Frequency", breaks = seq(min(chocolate$rating), max(chocolate$rating), length.out = 21))

## Histogram where bin num = 25
hist(chocolate$rating, main = "Overall distribution of rating scores (bin = 25)", xlab =  "Rating", ylab = "Frequency", breaks = 25)
```

The change that occurs when selecting a different number of bins is that the number of bars on the histogram change, when breaks = 10, theres 10 bars on the histogram across the x axis. When breaks = 15, theres 15 bars on the histogram across the x axis and so on. In my opinon, the one that looks the best is when breaks = 10. Typically, a histogram has all bars touching, and when we increase to bin = 15 , the space between and roughly 3.3 drops to 0 or so small we cannot see it visually. The same problem occurs in bin = 20. When we make it to bin = 25, there is so much gapping between the bins, the histogram takes the shape of a bar chart, and loses its shape completely. Additionally, throughout all bin varations, the negative skew (or left skew) remains clear, so we are able to draw relatively the same conclusions like mean < median < mode. Though bin = 15 had advanatges like providing more specific data, I think that what it adds does not outweigh the benefits of bin = 10. 

```{r}
#| label: Part 1: 2. Explore Data - Beans By Country 
#| warning: false

## Arranging Data by the country of origin 
arrange(chocolate, country_of_bean_origin)

## Grouping by the country of origin
by_country <- group_by(chocolate, country_of_bean_origin)

## Summarizing data based on country, and the sum of the count of ratings. 
country_and_review <- summarize(by_country, 
num_reviews = sum(!is.na(rating))
)
print(country_and_review, n =Inf)
```

The amount of reviews that come from each country of origin are shown in the output from the code above. 

```{r}
#| label: Part 1: Explore Data - 3. Ecuador
#| warning: false

## Filtering data for Ecuador 
only_Ecuador <- filter(chocolate, country_of_bean_origin == "Ecuador")

## Calc mean of ratings from new data frame 
avg_rating_ecu <- mean(only_Ecuador$rating)
avg_rating_ecu

## Print the average 
print(paste("The average rating of chocolate bars where the country of origin of beans is Ecuador is: ", avg_rating_ecu))

## Creating Ecuador Summary 
ecuador_summary <- only_Ecuador %>%
    summarize (
    total = sum(!is.na(rating)),
    mean = mean(rating),
    sd = sd(rating),
)
ecuador_summary
```


```{r}
#| label: Part 1: Explore Data - 4. Location with best chocolate 
#| warning: false

## Find location with Best Chocolate
best_location <- only_Ecuador %>%

## Group by the company location
  group_by(company_location) %>%
  
## Calculate the average rating 
  summarize(
    mean_rating = mean(rating, na.rm = TRUE),
  ) %>%
  
## Sort from highest average rating to lowest
  arrange(desc(mean_rating))

# Print locations
print(best_location, n  = Inf)
```

The company location with the best average ratings with beans from Ecuador is Australia. 

```{r}
#| label: Part 1: Explore Data - 5. Best across all origin 
#| warning: false

## 
## Find countries of origin with best average ratings
best_overall_ratings <- chocolate %>%

## Group by the country of origin 
  group_by(country_of_bean_origin) %>%
  
## Calculate the average rating 
  summarize(
    mean_rating = mean(rating, na.rm = TRUE),
  ) %>%
  
## Sort from highest average rating to lowest
  arrange(desc(mean_rating))

# Print locations
print(head(best_overall_ratings, 3))
```

The countries of bean origin with the best average ratings are 
Tobago with  3.62, China with 3.5, and Sao Tome & Principe with 3.5.

```{r}
#| label: Part 1: Explore Data - 6. remove w less than 10 reviews
#| warning: false

## Find countries of origin with best average ratings
best_ratings_10 <- chocolate %>%

## Group by the country of origin 
  group_by(country_of_bean_origin) %>%
  
## Calculate the average rating 
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),
    num_reviews = sum(!is.na(rating))
  ) %>%

## Filter out countries with less than 10 reviews 
filter(num_reviews >= 10) %>% 
  
## Sort from highest average rating to lowest
  arrange(desc(avg_rating))

# Print locations
print(head(best_ratings_10, 3))
```

Now, filtering for 10 reviews or more, the countries of bean origin with the best average ratings are Solomon Islands with 3.45 rating and 10 reviews, Congo with 3.32 rating and 11 reviews, and Cuba with 3.29 rating and 12 reviews. 

```{r}
#| label: Part 1: Explore Data - 7. Percent Chocolate and Ratings 
#| warning: false

## Identify the countries of bean origin with at least 50 reviews.
new_choco <- chocolate %>%

## Filter out countries with less than 50 reviews for country of origin 
  group_by(country_of_bean_origin) %>%
  filter(sum(!is.na(rating)) >= 50)
  
print(new_choco, n=Inf)

## Using mutate function to add a new column 
new_choco <- new_choco %>%
  mutate(

    ## Converting columns to numbers using substr and nchar - 1 for indexing str 
    cocoa_percent = as.numeric(substr(cocoa_percent, 1, nchar(cocoa_percent) - 1)),

    ## Using case when to break up mutate var 
    percent_groups = case_when(
      cocoa_percent < 60 ~ "<60%",
      cocoa_percent >= 60 & cocoa_percent < 70 ~ ">= 60% to <70%",
      cocoa_percent >= 70 & cocoa_percent < 90 ~ ">= 70% to <90%",
      cocoa_percent >= 90 ~ ">=90%"
    )
  )

## re-order factor levels (just in case and for practice).
new_choco <- new_choco %>%

  ## ungrouping to avoid error 
  ungroup()%>%

  ## implementing mutate and fct_relevel
  mutate(
    percent_groups = fct_relevel(percent_groups, "<60%", ">= 60% to <70%", ">= 70% to <90%", ">=90%")
  )


## Use ggplot to create boxplots, use aes to set x and y 
ggplot(new_choco, aes(x=percent_groups, y = rating)) + 
  ## Using geom.boxplot
  geom_boxplot() +

    ## this gets side by side plots (see source at end)
    facet_wrap( ~ country_of_bean_origin) + 

    ## (labels for graph)
    labs( x = "Percentage Groups", 
    y = "Rating",
    title = "Ratings by Chocolate Percentage Groups, Faceted by Country")


```

On average, which category of chocolate percentage is most highly rated? Do these countries mostly agree or are there disagreements?

On average, I think that ">= 60% to <70%" and the ">= 70% to <90%" groups are the groups that were the most highly rated, since both groups consistently have a high median category. It is hard to say because favor one group over the other, bur both are ranked on the higher end a majority of the time, one is not consistently higher than the other, and oftentimes they are tied. Countries tend to agree that the >=90% group rates consistently low.

Sources I used to help me answer this question: https://stackoverflow.com/questions/49458591/use-facet-wrap-in-r

I was was not familiar with this, so it was fun to learn about this way of using facet-wraps. I remember a different way of doing it from undergrad (I was a Statostics major, compsci minor), so this was actually quite interesting. 


