---
title: "Project 1"
author: "Grace Fischer"
date: "09/20/26"
format: html
---

# Setting Up
```{r}
#| label: access-data
#| warning: false

## Install the tidyverse package
if (!require("tidyverse", quietly = TRUE)) {
    install.packages("tidyverse")
}

library("tidyverse")

## Install the tidytuesdayR package
if (!require("tidytuesdayR", quietly = TRUE)) {
    install.packages("tidytuesdayR")
}

library("here")
library("tidyverse")

# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here("data", "chocolate.RDS"))) {
    url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-01-18/chocolate.csv"
    chocolate <- readr::read_csv(url_csv)

    # save the file to RDS objects
    saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}

# Read dataset locally from computing enviorment
chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)

# Glimpse at the data 
glimpse(chocolate)
```


# Part 1: Explore Data
```{r}
#| label: Part 1 - 1. Explore Data - Histogram
#| warning: false

## Histogram where bin num = 10 
hist(chocolate$rating, main = "Overall distribution of rating scores (bin = 10)", xlab =  "Rating", ylab = "Frequency", breaks = 10)

## Histogram where bin num = 15 
hist(chocolate$rating, main = "Overall distribution of rating scores (bin = 15)", xlab =  "Rating", ylab = "Frequency", breaks = 15)

## Histogram where bin num = 20 (breaks this way because r typically favors pretty numbers over round, length out is because it produces n-1 bins)
hist(chocolate$rating, main = "Overall distribution of rating scores (bin = 20)", xlab =  "Rating", ylab = "Frequency", breaks = seq(min(chocolate$rating), max(chocolate$rating), length.out = 21))

## Histogram where bin num = 25
hist(chocolate$rating, main = "Overall distribution of rating scores (bin = 25)", xlab =  "Rating", ylab = "Frequency", breaks = 25)
```

The change that occurs when selecting a different number of bins is that the number of bars on the histogram change, when breaks = 10, theres 10 bars on the histogram across the x axis. When breaks = 15, theres 15 bars on the histogram across the x axis and so on. In my opinon, the one that looks the best is when breaks = 10. Typically, a histogram has all bars touching, and when we increase to bin = 15 , the space between and roughly 3.3 drops to 0 or so small we cannot see it visually. The same problem occurs in bin = 20. When we make it to bin = 25, there is so much gapping between the bins, the histogram takes the shape of a bar chart, and loses its shape completely. Additionally, throughout all bin varations, the negative skew (or left skew) remains clear, so we are able to draw relatively the same conclusions like mean < median < mode. Though bin = 15 had advanatges like providing more specific data, I think that what it adds does not outweigh the benefits of bin = 10. 

```{r}
#| label: Part 1 - 2. Explore Data - Beans By Country 
#| warning: false

## Arranging Data by the country of origin 
arrange(chocolate, country_of_bean_origin)

## Grouping by the country of origin
by_country <- group_by(chocolate, country_of_bean_origin)

## Summarizing data based on country, and the sum of the count of ratings. 
country_and_review <- summarize(by_country, 
num_reviews = sum(!is.na(rating))
)
print(country_and_review, n =Inf)
```

The amount of reviews that come from each country of origin are shown in the output from the code above. 

```{r}
#| label: Part 1 - Explore Data - 3. Ecuador
#| warning: false

## Filtering data for Ecuador 
only_Ecuador <- filter(chocolate, country_of_bean_origin == "Ecuador")

## Calc mean of ratings from new data frame 
avg_rating_ecu <- mean(only_Ecuador$rating)
avg_rating_ecu

## Print the average 
print(paste("The average rating of chocolate bars where the country of origin of beans is Ecuador is: ", avg_rating_ecu))

## Creating Ecuador Summary 
ecuador_summary <- only_Ecuador %>%
    summarize (
    total = sum(!is.na(rating)),
    mean = mean(rating),
    sd = sd(rating),
)
ecuador_summary
```


```{r}
#| label: Part 1 - Explore Data - 4. Location with best chocolate 
#| warning: false

## Find location with Best Chocolate
best_location <- only_Ecuador %>%

## Group by the company location
  group_by(company_location) %>%
  
## Calculate the average rating 
  summarize(
    mean_rating = mean(rating, na.rm = TRUE),
  ) %>%
  
## Sort from highest average rating to lowest
  arrange(desc(mean_rating))

# Print locations
print(best_location, n  = Inf)
```

The company location with the best average ratings with beans from Ecuador is Australia. 

```{r}
#| label: Part 1 - Explore Data - 5. Best across all origin 
#| warning: false

## 
## Find countries of origin with best average ratings
best_overall_ratings <- chocolate %>%

## Group by the country of origin 
  group_by(country_of_bean_origin) %>%
  
## Calculate the average rating 
  summarize(
    mean_rating = mean(rating, na.rm = TRUE),
  ) %>%
  
## Sort from highest average rating to lowest
  arrange(desc(mean_rating))

# Print locations
print(head(best_overall_ratings, 3))
```

The countries of bean origin with the best average ratings are 
Tobago with  3.62, China with 3.5, and Sao Tome & Principe with 3.5.

```{r}
#| label: Part 1 - Explore Data - 6. remove w less than 10 reviews
#| warning: false

## Find countries of origin with best average ratings
best_ratings_10 <- chocolate %>%

## Group by the country of origin 
  group_by(country_of_bean_origin) %>%
  
## Calculate the average rating 
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),
    num_reviews = sum(!is.na(rating))
  ) %>%

## Filter out countries with less than 10 reviews 
filter(num_reviews >= 10) %>% 
  
## Sort from highest average rating to lowest
  arrange(desc(avg_rating))

# Print locations
print(head(best_ratings_10, 3))
```

Now, filtering for 10 reviews or more, the countries of bean origin with the best average ratings are Solomon Islands with 3.45 rating and 10 reviews, Congo with 3.32 rating and 11 reviews, and Cuba with 3.29 rating and 12 reviews. 

```{r}
#| label: Part 1 - Explore Data - 7. Percent Chocolate and Ratings 
#| warning: false

## Identify the countries of bean origin with at least 50 reviews.
new_choco <- chocolate %>%

## Filter out countries with less than 50 reviews for country of origin 
  group_by(country_of_bean_origin) %>%
  filter(sum(!is.na(rating)) >= 50)
  
print(new_choco, n=Inf)

## Using mutate function to add a new column 
new_choco <- new_choco %>%
  mutate(

    ## Converting columns to numbers using substr and nchar - 1 for indexing str 
    cocoa_percent = as.numeric(substr(cocoa_percent, 1, nchar(cocoa_percent) - 1)),

    ## Using case when to break up mutate var 
    percent_groups = case_when(
      cocoa_percent < 60 ~ "<60%",
      cocoa_percent >= 60 & cocoa_percent < 70 ~ ">= 60% to <70%",
      cocoa_percent >= 70 & cocoa_percent < 90 ~ ">= 70% to <90%",
      cocoa_percent >= 90 ~ ">=90%"
    )
  )

## re-order factor levels (just in case and for practice).
new_choco <- new_choco %>%

  ## ungrouping to avoid error 
  ungroup()%>%

  ## implementing mutate and fct_relevel
  mutate(
    percent_groups = fct_relevel(percent_groups, "<60%", ">= 60% to <70%", ">= 70% to <90%", ">=90%")
  )


## Use ggplot to create boxplots, use aes to set x and y 
ggplot(new_choco, aes(x=percent_groups, y = rating)) + 
  ## Using geom.boxplot
  geom_boxplot() +

    ## this gets side by side plots (see source at end)
    facet_wrap( ~ country_of_bean_origin) + 

    ## (labels for graph)
    labs( x = "Percentage Groups", 
    y = "Rating",
    title = "Ratings by Chocolate Percentage Groups, Faceted by Country")


```

On average, which category of chocolate percentage is most highly rated? Do these countries mostly agree or are there disagreements?

On average, I think that ">= 60% to <70%" and the ">= 70% to <90%" groups are the groups that were the most highly rated, since both groups consistently have a high median category. It is hard to say because favor one group over the other, bur both are ranked on the higher end a majority of the time, one is not consistently higher than the other, and oftentimes they are tied. Countries tend to agree that the >=90% group rates consistently low.

Sources I used to help me answer this question: https://stackoverflow.com/questions/49458591/use-facet-wrap-in-r

I was was not familiar with this, so it was fun to learn about this way of using facet-wraps. I remember a different way of doing it from undergrad (I was a Statostics major, compsci minor), so this was actually quite interesting. 

# Part 2: Join two datasets together
```{r}
#| label: Part 2 - Joining Data - 1. Join continents
#| warning: false

## Installing gapminder 
## install.packages("gapminder") already installed 
library("gapminder")

## Setting up continent before joining 
 country_by_cont <- gapminder::gapminder_unfiltered %>%
  distinct(country, continent)

## Left join to keep row in chocolate and add continent
chocolate <- chocolate %>%
  ## Making check for countries 
  left_join(country_by_cont, by = c("country_of_bean_origin" = "country"))

```

```{r}
#| label: Part 2 - Joining Data - 2. Bean Origin with at least 10 reviews 
#| warning: false

chocolate <- chocolate %>%

## Group by the country of origin 
  group_by(country_of_bean_origin) %>%
  
## Filter out countries with less than 10 reviews 
filter(sum(!is.na(rating)) >= 10)  %>%

## ungrouping to avoid error 
ungroup()
```

```{r}
#| label: Part 2 - Joining Data - 3. Removing "Blend"
#| warning: false

chocolate <- chocolate %>%

  ## Filter out countries with Blend
   filter(country_of_bean_origin != "Blend")

```

```{r} 
#| label: Part 2 - Joining Data - Naming NA countries
#| warning: false

## I waited to do this before graphs since we were removing countries with less than 10 reviews, and blend anyway! 

## Checking to see when a continent is NA 
  chocolate %>%
  filter(is.na(continent)) %>%
  distinct(country_of_bean_origin)

## Fixing NA continents listed 
chocolate <- chocolate %>%
  mutate(
    ## Using case when to prepare functions 
    continent = case_when(
      country_of_bean_origin %in% c("Sao Tome", "Congo") ~ "Africa",
      country_of_bean_origin %in% c("Trinidad", "St. Lucia", "U.S.A.") ~ "Americas",
      ## Need this for rows that already assigned
      TRUE ~ continent
    )
  )

## Run check again to make sure all countries have continents 
  chocolate %>%
  filter(is.na(continent)) %>%
  distinct(country_of_bean_origin)

```

```{r}
#| label: Part 2 - Joining Data - 4. Creating Violin Plot
#| warning: false

## Using ggplot and aes, setting x and y respectively 
ggplot(chocolate, aes(x = continent, y = rating)) + 
  geom_violin() + 
  labs(
    x = "Continent", 
    y = "Rating", 
    title = "Ratings by Continent")


```

# Part 3: Convert wide data into long data

```{r}
#| label: Part 3 - Wide to Long - 1. New Columns 
#| warning: false

chocolate <- chocolate %>%

## Accounting for NA's which would go undectected, mutate acts as boolean here and does the 0, 1's 
  mutate(
    beans        = replace_na(as.integer(str_detect(ingredients, "B")), 0),
    sugar        = replace_na(as.integer(str_detect(ingredients, "S,|S$")), 0),
    cocoa_butter = replace_na(as.integer(str_detect(ingredients, "C")), 0),
    vanilla      = replace_na(as.integer(str_detect(ingredients, "V")), 0),
    letchin      = replace_na(as.integer(str_detect(ingredients, "L")), 0),
    salt         = replace_na(as.integer(str_detect(ingredients, "Sa")), 0)
  )
```

```{r}
#| label: Part 3 - Wide to Long - 2. Memorable characteristics
#| warning: false


## Checking for ket words in most memorable characteritsics, here the muatate function acting as boolean, assign 0 is word not detected. 
chocolate <- chocolate %>%
  mutate(
    char_cocoa  = as.integer(str_detect(most_memorable_characteristics, "cocoa")),
    char_sweet  = as.integer(str_detect(most_memorable_characteristics, "sweet")),
    char_nutty  = as.integer(str_detect(most_memorable_characteristics, "nutty")),
    char_creamy = as.integer(str_detect(most_memorable_characteristics, "creamy")),
    char_roasty = as.integer(str_detect(most_memorable_characteristics, "roasty")),
    char_earthy = as.integer(str_detect(most_memorable_characteristics, "earthy"))
  )
```

```{r}
#| label: Part 3 - Wide to Long - 3. Mean across review Date
#| warning: false

## Calculate means for columns 
chocolate <- chocolate %>%
  ## Group by review year 
  group_by(review_date) %>%
  ## Summarize the means
  summarize(
    beans        = mean(beans),
    sugar        = mean(sugar),
    cocoa_butter = mean(cocoa_butter),
    vanilla      = mean(vanilla),
    letchin      = mean(letchin),
    salt         = mean(salt),
    char_cocoa   = mean(char_cocoa),
    char_sweet   = mean(char_sweet),
    char_nutty   = mean(char_nutty),
    char_creamy  = mean(char_creamy),
    char_roasty  = mean(char_roasty),
    char_earthy  = mean(char_earthy),
  )

## Check for specs: 16 rows 13 cols 
chocolate

```

```{r}
#| label: Part 3 - Wide to Long - 4. wide to long 
#| warning: false

## Sources I used to help me included at end
chocolate_long <- chocolate %>%
  pivot_longer(
    cols = beans:char_earthy,
    names_to = "feature",
    values_to = "mean_score"
  )

chocolate_long
```

Sources I used to help me complete this: https://stackoverflow.com/questions/72482406/how-to-use-pivot-longer-as-a-newbie 

# Part 4: Data Visualization 
#### Using chocolate_long from Part 3
```{r}
#| label: Part 4 - Data Visualization - Creating the scatterplot 
#| warning: false

## Use ggplot with data and aes for x and y respecively 
ggplot(chocolate_long, aes(x = review_date, y = mean_score)) +
  ## Plot type 
  geom_point() +
  ## 2. non-linear line (mean points above by geom.point)
  geom_smooth(method = "loess" ) +
  ## 3. all plots shown in 1 figure, with individual scales for numbers depending on data "free_y" (source below)
  facet_wrap(~ feature, scales = "free_y") +
  labs(
    ## 4. x and y axis labels 
    x = "Review Year",
    y = "Proportion of Reviews Mentioning Feature",
    ## 1. Title, subtitile, caption
    title = "Chocolate Ingreidents and Flavor Variations Through the Years", 
    subtitle = "Since around 2006, there has been a decline notes such as vanilla, lecithin, and cream, \nwhereas earthy, roasty, and cocoa notes have risen in popularity.", 
    caption = "Grace Fischer") +
  theme_light()


```

Sources I used to help me complete this: 
https://stackoverflow.com/questions/18046051/setting-individual-axis-limits-with-facet-wrap-and-scales-free-in-ggplot2

# Part 5: Make the worst plot you can
```{r}
#| label: Part 5 - Make the worst plot you can - 1. Making the Plot 
#| warning: false

## For this part, I will be using ggplot chocolate with all of the data from the data set, using aes to set x and y respectively

## 1. For my first modification I am going to choose the color of the data, based on which country the bean belong to, it will be colored accordingly.

## 2. My second modification is sizing by rating. This way, the scale of the dot will be based upon the rating value.
ggplot(new_choco, aes(x = country_of_bean_origin, y = rating, color = country_of_bean_origin, size = rating)) + 

  ## 3. The third modification is using geom_jitter() which offsets points so they are not overlapping (in this case since countries are the same, there would be some overlapping)
  ## 4. My fourth modification is using shape = 0 which makes the shape of the points a square that is hollow
  geom_jitter(shape = 0) + 
    ## 5. My fifth modification is fitting an lm line which is not necessarily good since country of origin is a categorical and not numerical variable
    ## 6. I also chose to make the linewidth 3, this way it is abnormally thick and draws attention away from data points. 
  geom_smooth(aes(group = 1), method = "lm", linewidth = 3) +

    ##7. I broke up the graphs by the percent_group scales and now 
    
    facet_wrap(~percent_groups, scales = "free" ) +
      ## 8. I added labels to the graphs and made them bad just to make the worst graph 
    labs(
    x = "country",
    y = "increasing nums rating",
    title = "Chocolate",
    subtitle = "wow, chocolate graph",
    caption = "Grace"
  ) + 
    ## 9. I made the plot background a different color 
    theme(
      plot.background = element_rect(fill = "gold"),
      panel.background = element_rect(fill = "chartreuse")
    )


```

Sources I used to help me complete this: https://ggplot2.tidyverse.org/reference/geom_jitter.html

1. Creating link between color to country_of_bean_origin here is not good because with so many countries it is hard to memorize, however, this could be good to use in in a visualization with a low number of categories, to keep the number of colors low and easy to remember, making it distinct. 
2. Sizing by rating in this graph is slightly redundant since rating is shown on y axis, however, in a good visualization using a different variable would make it more informative and useful. 
3. Using geom_jitter allows overlapping points to not be on top of one another so they can be seen better, but here there is a heavy seperation because of the data so it can misrepresent the data, making it look more spread out than it really is.  In a good visualization, when used carefully, it can be useful. 
4. Changing the shape of the points can be useful in a good visualization, like using high contrast shapes that are easy to read, even changing them to represent different categories. Here to make a bad graph I chose shapes that were intentionally difficult to read and would overlap because of aforemenitoned modifications, to prove my point. 
5. Fitting the lm line in this case makes the slope meaningless since it is categorical. In a good visualization, a trendline would fit the data and use loess or whatever shape would meet it. I had to add the aes line to make it show up. 
6. Making the line as thick as it is draws attention away from the data points where the line is suppose to enhance the data, not be a focal point. In a good graph, it would be subtle and help guide the viewer across the data, not distract. 
7. Faceting percent groups an using scale = free lets the axis range be independent on each of the panels. This makes it easy to misjudge, since they are all measuring the same thing. To use this in a good visualization, all scales would be the same or scale = free would be used when the range is unrelated to the others. 
8. This is not much of a modification but i felt it should be included. The use of labels, titiles, subtitiles, and captions should be meaningful in a good visualization so the viewer can figure out what is going on. 
9. Using panel and plot colors that are clashing is an eye-sore and distracting. In a good visualization, we can use differeny colors to make it interesting, but still neutral and low contrast so that the data remains the focal point, and easy to read. 


# Part 6: Make my plot a better plot!
```{r}
#| label: Part 6 - Make my plot a better plot!
#| warning: false

library("tidyverse")
chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)

## Original plot
chocolate %>%
    ggplot(aes(
        x = as.factor(review_date),
        y = rating,
        fill = review_date
    )) +
    geom_violin()

## Improved Plot
chocolate %>%
    ggplot(aes(
        x = as.factor(review_date),
        y = rating,
    )) +
      ## 1. Using one color, the color distinction does not add anything, as year is already expressed on x axis, it is just redunant.
      ## 2. alpha to color so its not as opaque 
    geom_violin(fill = "lightblue", color = "black", alpha = 0.7) +
    ## 3. Added stat summary to show where Q1, median and Q3 are. I added source below. 
    stat_summary(
      fun.min = function(x) quantile(x, 0.25),
      fun.max = function(x) quantile(x, 0.75),
      fun = median,
      geom = "pointrange"
      ) + 
        ## 4. Added this since points go 1-4, helps with white space 
        coord_cartesian(ylim = c(1, 4)) +
      ## 5. Added labs to make graph more specific 
    labs(
    x = "Year of Review",
    y = "Rating",
    title = "Chocolate Bar Ratings by Year",
    subtitle = "Rating distributions have remained fairly constant, with a slight upward\nshift as time goes on"
    ) + 

    ## 6. Added a theme for consistensy and ease of reading. 
    ## 7. Added base font 
    theme_minimal(base_family = "Georgia") +
    ## 8. Adjusted fonts, angle, etc for clear reading, source below 
    theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(color = "darkgray", size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1)
    )
```

Sources I used to help me complete this: 
https://stackoverflow.com/questions/41077199/how-to-use-r-ggplot-stat-summary-to-plot-median-and-quartiles

https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2 

1. I removed the fill = review date. In the original plot it was displaying a variable that was already being shown on the plot. It displaying an unnecessary gradient that did not provide any additional information to the viewer, so instead I made it a single color instead of the gradient with a key. 
2. I added alpha to the fill. After doing teh fill I thought the colors I was trying all came out too opaque and heavy. This adds softness and an element of transparency to the graph, overall it makes it look more polished. 
3. I added stat_summary. Here, it displays Q1, Q2 (median), and Q3. Though the violin shows the overall distribution shape, it has no summary or specific statistics. I felt that this gave the viewer the exact IQR and median qithout the entire boxplot making it look to busy or heavy.
4. I used coord_cartesian to do y-axis range. It only goes from 1-4, so this removes any uncessary white space. Though there was not much to begin with, incase there would be after any edits this got tid of that and also keeps violin shapes and quartile calculations accurate.
5. I added labels to the graph to help the viewer better understand what they are looking at. The original plot had no labels at all, and even left the x label as what it is in the code. Adding this helps the viewer understand what they are looking at. 
6. I added the minimal theme. The default theme has a heavy visual weight compared to the violin graphs, this filters out some of the background additions it had before. 
7. I changed the font to Georgia. The font before felt plain and changing to a serif font felt more finished and clean to me 
8. I customized the bold, color, and angle of the labels. Thsi way the viewer can distinguish whagt to read first. Additionally, angling the years prevents them from overlapping one another and potentionally being unreadable at any point.

